---
title: "Gov 51: Problem Set 1"
author: "Amy Tan"
format: pdf
editor: visual
---

## Question 1.7

Github link: <https://github.com/amyqtan/gov51-ps1>

## Question 2.1

```{r}
library(tidyverse)
acs2024 <- read_csv("/Users/amytan/Desktop/gov51-ps1/data/raw/acs2024.csv")
```

## Question 2.2

TRANTIME is a 3-digit numeric variable reporting the total amount of time, in minutes, that it usually took the respondent to get from home to work last week. A zero value for TRANTIME would represent a zero minute commute time, implying that the person either works from home, does not work, etc. Fundamentally, a zero minute commute time means that the respondent is not a commuter.

PERWT is a 6-digit numeric variable which indicates how many persons in the U.S. population are represented by a given person in an IPUMS sample, and it has two implied decimals (meaning that the last two digits of the value are decimals, and the third-to-last digit is the start of the ones place. This helps later, for weighted means, since it will allow us to weight the sample to be nationally representative.

## Question 2.3

```{r}
acs2024_new <- acs2024 %>% mutate(
  female = case_when(SEX == 2 ~ 1, TRUE ~ 0),
  less_than_hs = case_when(EDUC %in% c(0, 1, 2, 3, 4 ,5) ~ 1, TRUE ~ 0),
  hs_only = case_when(EDUC == 6 ~ 1, TRUE ~ 0),
  some_college = case_when(EDUC %in% c(7, 8, 9) ~ 1, TRUE ~ 0),
  college_only = case_when(EDUC == 10 ~ 1, TRUE ~ 0),
  advanced_degree = case_when(EDUC == 11 ~ 1, TRUE ~ 0),
  employed = case_when(EMPSTAT == 1 ~ 1, TRUE ~ 0),
  unemployed = case_when(EMPSTAT == 2~ 1, TRUE ~ 0),
  not_labor_force = case_when(EMPSTAT == 3 ~ 1, TRUE ~ 0))
```

## Question 2.4

```{r}
acs2024_new_clean <- acs2024_new %>% mutate(INCTOT = na_if(INCTOT, 9999999),
                                            INCTOT = na_if(INCTOT, 9999998))
```

Accidentally treating N/A codes as real zeroes could cause artificially low mean values, since those values are not real zeroes and are simply unknown. A solution to this could be to impute the data.

For instance, for the variable `INCTOT`, the value 9999999 means the total income is N/A and the value 9999998 means the total income is unknown. If these were treated as actual values in the dataset, it would artificially raise the income total mean, since these are not actual total income values but rather unknown and N/A values. If they are treated as zeroes instead of N/A values, this would inaccurately lower the mean.

## Question 2.5 (summary table)

```{r}
acs2024_summarytable <- data.frame(
  Variable = c("Age", "Female", "Less than High School", 
               "High School Only", "Some College", "College Only", 
               "Advanced Degree", "Employed", "Unemployed", 
               "Not in Labor Force", "Commute Time (mins)", 
               "Total Income ($)"),
  N = c(length(acs2024_new$AGE),
        length(acs2024_new$female),
        length(acs2024_new$less_than_hs),
        length(acs2024_new$hs_only),
        length(acs2024_new$some_college),
        length(acs2024_new$college_only),
        length(acs2024_new$advanced_degree),
        length(acs2024_new$employed),
        length(acs2024_new$unemployed),
        length(acs2024_new$not_labor_force),
        length(acs2024_new$TRANTIME),
        length(acs2024_new$INCTOT)),
  Mean = c(mean(acs2024_new$AGE),
           mean(acs2024_new$female),
           mean(acs2024_new$less_than_hs),
           mean(acs2024_new$hs_only),
           mean(acs2024_new$some_college),
           mean(acs2024_new$college_only),
           mean(acs2024_new$advanced_degree),
           mean(acs2024_new$employed),
           mean(acs2024_new$unemployed),
           mean(acs2024_new$not_labor_force),
           mean(acs2024_new$TRANTIME),
           mean(acs2024_new$INCTOT)),
  Std_Dev = c(sd(acs2024_new$AGE),
           sd(acs2024_new$female),
           sd(acs2024_new$less_than_hs),
           sd(acs2024_new$hs_only),
           sd(acs2024_new$some_college),
           sd(acs2024_new$college_only),
           sd(acs2024_new$advanced_degree),
           sd(acs2024_new$employed),
           sd(acs2024_new$unemployed),
           sd(acs2024_new$not_labor_force),
           sd(acs2024_new$TRANTIME),
           sd(acs2024_new$INCTOT)),
  Min = c(min(acs2024_new$AGE),
           min(acs2024_new$female),
           min(acs2024_new$less_than_hs),
           min(acs2024_new$hs_only),
           min(acs2024_new$some_college),
           min(acs2024_new$college_only),
           min(acs2024_new$advanced_degree),
           min(acs2024_new$employed),
           min(acs2024_new$unemployed),
           min(acs2024_new$not_labor_force),
           min(acs2024_new$TRANTIME),
           min(acs2024_new$INCTOT)),
  Max = c(max(acs2024_new$AGE),
           max(acs2024_new$female),
           max(acs2024_new$less_than_hs),
           max(acs2024_new$hs_only),
           max(acs2024_new$some_college),
           max(acs2024_new$college_only),
           max(acs2024_new$advanced_degree),
           max(acs2024_new$employed),
           max(acs2024_new$unemployed),
           max(acs2024_new$not_labor_force),
           max(acs2024_new$TRANTIME),
           max(acs2024_new$INCTOT)))

```

\newpage

```{r}
knitr::kable(acs2024_summarytable, 
             digits = 2, 
             caption = "Summary Statistics for 2024 ACS Sample (Before Marking NA Values)") 

```

As seen in this table, treating the NA and unknown values of 9999999 and 9999998 artificially raised the mean of total income, making it \$1,536,765.79, which is not a realistic mean income for the United States. This shows why it is important to accurately treat N/A values as N/As and not zeroes or as the numbers themselves. The next table properly treats the N/A values as N/A.

```{r}
acs2024_summarytable_clean <- data.frame(
  Variable = c("Age", "Female", "Less than High School", "High School Only", 
               "Some College", "College Only", "Advanced Degree", 
               "Employed", "Unemployed", "Not in Labor Force", 
               "Commute Time (mins)", "Total Income ($)"),
  N = c(length(acs2024_new_clean$AGE),
        length(acs2024_new_clean$female),
        length(acs2024_new_clean$less_than_hs),
        length(acs2024_new_clean$hs_only),
        length(acs2024_new_clean$some_college),
        length(acs2024_new_clean$college_only),
        length(acs2024_new_clean$advanced_degree),
        length(acs2024_new_clean$employed),
        length(acs2024_new_clean$unemployed),
        length(acs2024_new_clean$not_labor_force),
        length(acs2024_new_clean$TRANTIME),
        sum(!is.na(acs2024_new_clean$INCTOT))),
  Mean = c(mean(acs2024_new_clean$AGE),
           mean(acs2024_new_clean$female),
           mean(acs2024_new_clean$less_than_hs),
           mean(acs2024_new_clean$hs_only),
           mean(acs2024_new_clean$some_college),
           mean(acs2024_new_clean$college_only),
           mean(acs2024_new_clean$advanced_degree),
           mean(acs2024_new_clean$employed),
           mean(acs2024_new_clean$unemployed),
           mean(acs2024_new_clean$not_labor_force),
           mean(acs2024_new_clean$TRANTIME),
           mean(acs2024_new_clean$INCTOT, na.rm = TRUE)),
  Std_Dev = c(sd(acs2024_new_clean$AGE),
           sd(acs2024_new_clean$female),
           sd(acs2024_new_clean$less_than_hs),
           sd(acs2024_new_clean$hs_only),
           sd(acs2024_new_clean$some_college),
           sd(acs2024_new_clean$college_only),
           sd(acs2024_new_clean$advanced_degree),
           sd(acs2024_new_clean$employed),
           sd(acs2024_new_clean$unemployed),
           sd(acs2024_new_clean$not_labor_force),
           sd(acs2024_new_clean$TRANTIME),
           sd(acs2024_new_clean$INCTOT, na.rm = TRUE)),
  Min = c(min(acs2024_new_clean$AGE),
           min(acs2024_new_clean$female),
           min(acs2024_new_clean$less_than_hs),
           min(acs2024_new_clean$hs_only),
           min(acs2024_new_clean$some_college),
           min(acs2024_new_clean$college_only),
           min(acs2024_new_clean$advanced_degree),
           min(acs2024_new_clean$employed),
           min(acs2024_new_clean$unemployed),
           min(acs2024_new_clean$not_labor_force),
           min(acs2024_new_clean$TRANTIME),
           min(acs2024_new_clean$INCTOT, na.rm = TRUE)),
  Max = c(max(acs2024_new_clean$AGE),
           max(acs2024_new_clean$female),
           max(acs2024_new_clean$less_than_hs),
           max(acs2024_new_clean$hs_only),
           max(acs2024_new_clean$some_college),
           max(acs2024_new_clean$college_only),
           max(acs2024_new_clean$advanced_degree),
           max(acs2024_new_clean$employed),
           max(acs2024_new_clean$unemployed),
           max(acs2024_new_clean$not_labor_force),
           max(acs2024_new_clean$TRANTIME),
           max(acs2024_new_clean$INCTOT, na.rm = TRUE)))
```

\newpage

```{r}
knitr::kable(acs2024_summarytable_clean, 
             digits = 2, 
             caption = "Summary Statistics for 2024 ACS Sample (After Marking NA Values)") 

```

## Question 2.6

The commute time of 195 minutes (3 hours and 15 minutes) makes sense for a commute time. While this is an extremely long commute, there are some people who make extremely long commutes in America because they work in extremely high-cost of living areas (and therefore cannot afford to live near where they work) or because they work extremely specialized jobs.

## Question 3.1

```{r}
#number of commuters with a commute time of zero
zero_trantime <- sum(acs2024_new$TRANTIME == 0) 
zero_trantime

#total number of respondents

total_respondents <- acs2024_new |> nrow() 
total_respondents

#percentage of respondents with a commute time of zero
zero_trantime / total_respondents * 100 
```

According to the data, 2,062,945 people have a commute time of zero. This is 60.27% of total commuters (calculated by taking the number of people with zero commute time and dividing by the total number of respondents and then using both of those values to calculate a percentage).

The people with zero commute time are likely remote workers, unemployed people, children, retirees, and people who do not commute at all because they don't have a commute.

## Question 3.2

```{r}
commuters <- acs2024_new |> filter(TRANTIME > 0)
```

Commuters filters the data into a subset of people who have a commute time greater than zero. These are people who actually have to commute to work, as opposed to the people who have a value of zero for their commute time, who should not be counted as commuters, because they are not commuting at all.

## Question 3.3

```{r}
total_commuters <- commuters |> nrow() #total number of commuters
total_commuters
total_respondents - zero_trantime
```

1,359,943 observations remain in the new dataset. To verify this was coded correctly, I subtracted the total number of respondents from those who had zero commute time (both of these variables were coded in question 3.1). This value should be equal to the number of rows in the new commuters dataset, since the new dataset filters out respondents with zero commute time from the total respondents.

The new minimum TRANTIME is 1 minute, which makes sense, because that is a commute (albeit very short but possible, such as living next door to the workplace) as shown in the table in question 3.4.

## Question 3.4

```{r}
commuters_summarytable <- data.frame(
  Variable = c("Age", "Female", "Less than High School", "High School Only", "Some College", "College Only", "Advanced Degree", "Employed", "Unemployed", "Not Labor Force", "Commute Time (mins)", "Total Income ($)"),
  N = c(length(commuters$AGE),
        length(commuters$female),
        length(commuters$less_than_hs),
        length(commuters$hs_only),
        length(commuters$some_college),
        length(commuters$college_only),
        length(commuters$advanced_degree),
        length(commuters$employed),
        length(commuters$unemployed),
        length(commuters$not_labor_force),
        length(commuters$TRANTIME),
        length(commuters$INCTOT)),
  Mean = c(mean(commuters$AGE),
           mean(commuters$female),
           mean(commuters$less_than_hs),
           mean(commuters$hs_only),
           mean(commuters$some_college),
           mean(commuters$college_only),
           mean(commuters$advanced_degree),
           mean(commuters$employed),
           mean(commuters$unemployed),
           mean(commuters$not_labor_force),
           mean(commuters$TRANTIME),
           mean(commuters$INCTOT)),
  Std_Dev = c(sd(commuters$AGE),
           sd(commuters$female),
           sd(commuters$less_than_hs),
           sd(commuters$hs_only),
           sd(commuters$some_college),
           sd(commuters$college_only),
           sd(commuters$advanced_degree),
           sd(commuters$employed),
           sd(commuters$unemployed),
           sd(commuters$not_labor_force),
           sd(commuters$TRANTIME),
           sd(commuters$INCTOT)),
  Min = c(min(commuters$AGE),
           min(commuters$female),
           min(commuters$less_than_hs),
           min(commuters$hs_only),
           min(commuters$some_college),
           min(commuters$college_only),
           min(commuters$advanced_degree),
           min(commuters$employed),
           min(commuters$unemployed),
           min(commuters$not_labor_force),
           min(commuters$TRANTIME),
           min(commuters$INCTOT)),
  Max = c(max(commuters$AGE),
           max(commuters$female),
           max(commuters$less_than_hs),
           max(commuters$hs_only),
           max(commuters$some_college),
           max(commuters$college_only),
           max(commuters$advanced_degree),
           max(commuters$employed),
           max(commuters$unemployed),
           max(commuters$not_labor_force),
           max(commuters$TRANTIME),
           max(commuters$INCTOT)))
```

\newpage

```{r}
knitr::kable(commuters_summarytable, 
             digits = 2, 
             caption = "Summary Statistics for only Commuters from 2024 ACS Sample") 
```

## Question 4.1

```{r}
library(ggplot2)
commuters_graph <- ggplot(data = commuters, mapping = aes(x = TRANTIME)) + 
  geom_histogram(binwidth=5, boundary = 0, fill = "blue", color = "white") + 
  labs(
    x = "Commute Time", 
    y = "Number of Commuters", 
    title = "Distribution of Commute Times in 2024 ACS Sample Data (Commuters Only)") + 
  scale_x_continuous(breaks = seq(0, 200, by = 10))
commuters_graph
```

## Question 4.2

The histogram is significantly right skewed, and the center appears to be in the range of 10 - 15 minutes of commute time, closer to 15 mintues. Most commute times seem to be in the range of 0 - 30 minutes, but there is a notable tail into commute times of 90 minutes from this cluster. Additionally, there is a very small outlier of commute times between 140-150 minutes.

## Question 4.3

```{r}
median(commuters$TRANTIME)
mean(commuters$TRANTIME)
```

The mean commute time of 27.22 minutes is larger than the median of 20 minutes, meaning that the histogram should be right-skewed. This matches what is seen in the histogram.

## Question 4.4

```{r}
ggsave("/Users/amytan/Desktop/gov51-ps1/output/commute_histogram.png", plot = commuters_graph)
```

## Question 5.1

```{r}
weighted.mean(commuters$TRANTIME, commuters$PERWT)
```

The weighted mean is 27.19 minutes.

## Question 5.2

```{r}
mean(commuters$TRANTIME)
mean(commuters$TRANTIME) - weighted.mean(commuters$TRANTIME, commuters$PERWT)
```

The unweighted mean is 27.22 minutes. There is a difference of 0.024 between the unweighted and weighted mean.

## Question 5.3

The unweighted mean represents the mean commute time within the sample of respondents in commuters. The weighted mean represents a mean commute time of the sample that is adjusted to be nationally representative. The weighted mean seeks to correct for oversampling and adjusts how many people in the U.S. population are represented by a given respondent within the survey data. They differ because while the unweighted mean treats each respondent as the same in the mean calculation, the weighted mean assigns a certain number of people in the United States for each respondent to represent (in order to make the mean nationally representative), so each respondent is not worth the same in the weighted mean.
